---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

#Peer Assessment 1

```{r setoptions, echo=F}
library(knitr)
opts_chunk$set(echo=TRUE)
```

###Loading and preprocessing the data
The variables included in this dataset are:  
steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)  
date: The date on which the measurement was taken in YYYY-MM-DD format  
interval: Identifier for the 5-minute interval in which measurement was taken
```{r}
if (!file.exists("activity.csv")) {
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip")
        unzip("activity.zip")
}
data<-read.csv("activity.csv",sep=",")
```

###What is mean total number of steps taken per day?
1.First we calculate the total number of steps taken per day
```{r}
tspd<-tapply(data$steps,data$date,sum)
#we ignore the missing values
tspd<-tspd[complete.cases(tspd)]
tspd
```  
2.We make a histogram of the total number of steps taken each day
```{r}
n.clases<-7
points<-min(tspd)+(0:n.clases)*(max(tspd)-min(tspd))/n.clases
hist(tspd,breaks=points,col="lightblue",main="Histogram total number of steps per day",xlab="number of steps")
```  

3.We calculate the mean and median of the total number of steps taken per day
```{r}
mean(tspd)
median(tspd)
```  
###What is the average daily activity pattern?
1. We make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
data1<-data[complete.cases(data),]
asbi<-tapply(data1$steps,as.factor(data1$interval),mean)
plot(unique(data1$interval),asbi,type="l",main="Average number of steps",xlab="5 minutes intervals",ylab="Average of steps")
```  

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
interval<-row.names(asbi)
df<-data.frame(interval,asbi)
int<-df[df$asbi==max(asbi),]$interval
```

The maximum number of steps are in the interval **`r int`**  

###Imputing missing values

1. Calculate and report the total number of missing values in the dataset
```{r}
sum(!complete.cases(data))
```  
2. We fill the missing values using the mean for that interval
3. We create a new dataset filling the missing values:
```{r}
n.ind<-dim(data)[1]
newdata<-data
for(i in 1:n.ind){
    if(is.na(newdata[i,]$steps)){
        newdata[i,]$steps<-df[df$interval==data[i,]$interval,]$asbi
    }else{
        newdata[i,]$steps<-data[i,]$steps
    }
}
```

4. Make a histogram of the total number of steps taken each day.
```{r}
tspd2<-tapply(newdata$steps,newdata$date,sum)
n.clases<-7
points<-min(tspd2)+(0:n.clases)*(max(tspd2)-min(tspd2))/n.clases
hist(tspd2,breaks=points,col="lightblue",main="Histogram total number of steps per day",xlab="number of steps")
```

Calculate and report the mean and median total number of steps taken per day:
```{r}
mean(tspd2)
median(tspd2)
```

The only value that differ from from the estimates from the first part of the assignment is the median but little.

###Are there differences in activity patterns between weekdays and weekends?

1. We create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
newdata$date<-as.Date(newdata$date)
weekdays<-weekdays(newdata$date)
weekdays<-ifelse(weekdays=="sábado"|weekdays=="domingo","weekend","weekday")
newdata<-data.frame(newdata,weekdays)
```
2. We make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
library(lattice)
splitdata<-aggregate(steps~interval+weekdays,newdata,FUN="mean")
xyplot(steps~interval|weekdays,data=splitdata,type="l",main="Average number of steps in weekday vs weekend",ylab="Number of steps",xlab="5 minutes intervals",layout=c(1,2))
```


